---
title: "Class_Notes"
author: "Alejandra Barrera"
date: "6/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = ">")
```

# **Data Science - Coursera**  
Notas para la futura Alejandra. Recuerda **SIEMPRE** estos pasos: Guardar->Commit->Push (asi se hayan desaparecido los archivos) y ya. Puedes usar esta pagina donde estan las referencias de R Markdown.  

[Markdown Sheets](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)  
[Markdown Info](https://bookdown.org/yihui/rmarkdown/r-code.html)  
[R Cheat Sheets](http://datasciencefree.com/basicR.pdf) 

> Para borrar la consola cntrl+L y para borrar todas las variables rm(list=ls()). Para tener informacion de una funcion se pone _"?funcion"_    

There are different types of classes of objetcs. R has five basic or atomic classes of objects:  

  - character  
  - numeric  
  - integer (To specify the use `L` something like 5L) 
  - complex  
  - Logical  

R objects can have attributes which are like metadata. Attributes can be accessed using `attributes()`
        
  - names, dimnames  
  - dimensions  
  - class  
  - length  
  - other  
  
### __*Vectores, Matrices, listas, factores, dataframe...*__  {.tabset .tabset-pills}
#### Intro
  * Functions  
    + `c()`
    + `vector()`  
    + `names()`
    + `class()`  
    + `dimnames()`
    + `list()`
    + `factors()`  
    + `levels()` 
    + `is.na()`/ `is.nan()`  

#### Vectors  
Vectors can have only one class of objects, when different objects are mixed in a vector,
coercion occurs so that every element in the vector is of the same class. To create a vector:  

  * Con la funcion c()  
```{r}
x<- c(0.5, 0.6) #Como si c significara concatenar
print(x)
```

  * Tambien se le pueden poner nombres a las columnas  
```{r}
names(x)<-c("Food","Bar")
x
```
Pueden haber distintos objetos (numericos, logicos, enteros, etc)

  * Con la funcion vector() 

```{r}
x<-vector("numeric", length=10)
x
```
Also, different modes can be used such as "logical", "integer", "numeric" (synonym "double"), "complex", "character" and "raw".  
***
Para saber la clase entonces se usa `class()`
```{r}
class(x)
```
  * Se puede forzar la conversion entre clases de numeros. Por ejemplo:
```{r}
x<-0:6
class(x)
as.numeric(x)
as.character(x)
as.logical(x)
```
  * Vectors dont have dimentions. However, can be assigned:  
```{r}
my_vector <-1:20  
dim(my_vector)
dim(my_vector) <- c(4,5) #rows and cols
my_vector # now is a matrix
```
 
#### Lists
Are a special type of vectors that contain elements of different classes. To create one:  

  * Using `list()`  
  
```{r}
x<-list(1, "a", TRUE, 1+4i) #Recuerda que los numeros que cambian son los indices de los vectores, tienen dobles corchetes cuadrados
print(x)
```

  * Create an empty list with `vector()`  
```{r}
x <- vector("list", length = 5)
```
Tambien pueden tener nombres
```{r}
x<-list(a=1,b= "a",c= TRUE,d= 1+4i)
x
```
Para eliminar una lista se puede usar `unlist()`  

  * To access a list $ is used  
```{r}
x$b
```

#### Matrices
Matrices are vectores with a _dimmension_ attribute. Se construyen _column wise_ es decir, se llenan las filas de arriba para abajo una columna a la vez.
```{r}
m<- matrix(1:6, nrow=2, ncol=3) # Se construye column wise
m
dim(m)
attributes(m) #Todo lo que se le pueda sacar
m[1,] #Para extraer solo la primera fila

```
  * Pueden tener nombres. Also `rownames()` or `colnames()` can be used  
```{r}
dimnames(m) <- list(c("fila1","fila2"),c("col1","col2","col3")) #Se usa una listaaa
m
```

 * Otra forma es redimensionar un vector
```{r}
m<-1:10
dim(m) <- c(2,5) #dos filas, dos columnas
m
```
 * Cbind-ing and rbind-ing
```{r}
x<-1:3
y<-10:12
cbind(x,y) # Los une por columnas
print("Tambien se puede unir por filas")
rbind(x,y)
```

#### Factors  
One can think of a factor as an integer vector where each integer has a label. Factors are automatically created when `read.table()` is used.
Para tener datos categoricos:
```{r}
x<-factor(c("yes","no","yes", "no"), levels = c("yes","no")) #Para determinar el orden de los niveles.
x
table(x) #Conteo
class(x)
```

#### Data Frame  
Are used to stored tabular data in R. `dplyr` is a package optimized for data frames. Are represented as a special type of list where every element of the list has to have the same length. Can store different classes. Generally are read using functions such as `read.table()` or `read.csv()`
```{r}
x<-data.frame(food=1:4, bar=c(T,T,F,F))
x
nrow(x)
ncol(x)
```
  * Para cambiar el nombre de las columnas. `names()` can be used tu columns
```{r}
colnames(x)[c(1,2)]<-c("Name","Mortality"); x
x<-data.frame(food=1:4, bar=c(T,T,F,F))
names(x) <- c("Name","Mortality")
```
To set the names in the rows use `row.names()`  

#### Missing values  
Missing values are denoted by NA or NaN for undefined mathematical operations.  

  * `is.na()` is used to test objects if they are NA
  * `is.nan()` is used to test for NaN
  * NA values have a class also, so there are integer NA, character NA, etc.
  * NaN value is also NA but the converse is not true  
    
### __*Subsetting*__  {.tabset .tabset-pills}
#### Intro  
  * The [ operator always returns an object of the same class as the original. It can be used to select
multiple elements of an object  
* The [[ operator is used to extract elements of a list or a data frame. It can only be used to
extract a single element and the class of the returned object will not necessarily be a list or
data frame.  
* The $ operator is used to extract elements of a list or data frame by literal name. Its semantics
are similar to that of [[.  

* Functions  
    + `c()`
    + `vector()`  
    + `names()`
    + `class()`  
    + `dimnames()`
    + `list()`
    + `factors()`  
    + `levels()` 
    + `is.na()`/ `is.nan()`  
    + `identical()`
    
#### Vector 
```{r}
x <- c("a", "b", "c", "c", "d", "a")
x[1] ## Extract the first element
x[c(1, 3, 4)] # The sequences doesnt has to be in order
x[-c(1, 3, 4)] # All elements but those
```
We can also pass a logical sequence to the [ operator to extract elements of a vector that satisfy a
given condition. For example, here we want the elements of x that come lexicographically after the
letter “a”.  

```{r}
x[x > "a"]
```
  * Compare if two vectors are the same  
```{r}
identical(x,c(1,2))
```

#### Matrix  
```{r}
x <- matrix(1:6, 2, 3)
x[1, 2]
x[1, , drop = FALSE] #if the answer has to be a matrix
```
#### Lists  
   * Normal lists  
```{r}
x <- list(foo=1:4, bar=0.6) #lista
x[1] #Lista
x[[1]]
class(x[2])
class(x[[2]])
x$foo
x[[1]][[2]] #extraer un elemento de la lista
```
  * Nested Lists  
```{r}
x <- list(a = list(10, 12, 14), b = c(3.14, 2.81)); x
## Get the 3rd element of the 1st element
x[[c(1, 3)]] #or x[[1]][[3]]
```
    
### __*Estructuras de control*__ {.tabset .tabset-pills}
#### Intro
  * if and else: testing a condition and acting on it  
  * for: execute a loop a fixed number of times  
  * while: execute a loop while a condition is true  
  * repeat: execute an infinite loop (must break out of it to stop)  
  * break: break the execution of a loop  
  * next: skip an interation of a loop  

#### If Else  
```{r}
condition<-TRUE
if(condition==TRUE) {
  ## do something
  print(TRUE)
} else #else if 
  { print(FALSE)
  ## do something else
}
```
#### for  
Se puede usar _next_ para cambiar de iteracion o para salir.
```{r}
for(i in 1:10){
  print(i)
}
## otra forma de escribirlo
for(i in 1:4) print(i)
## Para crear una sequencia
seq(5)
```

#### While  
Se puede reemplazar con un _repeat_ pero hay que usar un _break_
```{r}
count<-0
while(count<10) {
  print(count)
  count<-count+1
}
```

### __*Funciones utiles*__ {.tabset .tabset-pills}
#### Intro  
To see the source code , just type the function without parenthesis. To see the arguments type `args(function)`  

  * `seq()`  
  *  `seq_along()` 
  * Funciones logicas
    + `isTRUE()`
    + `xor()`
    + `all()`
    + `any()`  
  * `which()`  
  
#### Funciones Anonimas  
```{r}
evaluate <- function(func, dat){func(dat)}
evaluate(function(x){x[1]}, c(8, 4, 0)) #return the first element of the vector
evaluate(function(x){x[length(x)]}, c(8, 4, 0))  #return the last element of the vector
```

#### Sequences  
How to create a sequence
```{r}
seq(1,10, by=2)
seq(1,10,length=3)
seq(along=c(1,3,8,25,100)) #create a vector with the length of the vector
seq_along(c(1,3,8,25,100)) #Same

```
#### Funciones logicas
The & is first evaluated than |. && or || evaluates only the first argument.
  * `isTRUE()`
```{r}
isTRUE(3>2)
```

  * The `xor()` function stands for exclusive OR. If one argument evaluates to TRUE and one argument evaluates to FALSE, then this function will return TRUE, otherwise it will return FALSE.  
```{r}
xor(5==6, !FALSE)
xor(TRUE, TRUE)
```
  * The any() function will return TRUE if one or more of the elements in the logical vector is TRUE. The all() function will return TRUE if every element in the logical vector is TRUE.
```{r}
ints <- sample(10)
any(ints<0)
all(ints>0)
```

#### Index  
  * Para encontrar los indices se usa `which()` 
```{r}
ints <-sample(10)
which(ints>5)
```
### __*Looping functions*__ {.tabset .tabset-pills}
#### Intro
  * `lapply()`
  * `sapply()`
  * `apply()`
  * `mapply()` `vectorize()`
  * `tapply()`
  * `split()`
  
#### lapply
The lapply() function does the following simple series of operations:  
  1. it loops over a list, iterating over each element in that list  
  2. it applies a function to each element of the list (a function that you specify)  
  3. and returns a list (the l is for “list”).    

```{r}
x<-list(a=matrix(1:4, 2,2), b=matrix(1:6,3,2)); x
lapply(x, mean)
```
When you pass a function to lapply(), lapply() takes elements of the list and passes them as the
first argument of the function you are applying. 

Si se tiene una lista, se va a aplicar la funcion a toda la lista.  
```{r}
library(datasets)
data(iris)
iris[1:5,]
lapply(iris,mean)
lapply(x, function(elt) { elt[,1] }) #Elt is the argument, i.e the list

```
You can put an arbitrarily complicated function definition inside lapply(), but if it’s
going to be more complicated, it’s probably a better idea to define the function separately.

#### sapply 
The sapply() function behaves similarly to lapply(); the only real difference is in the return value.
sapply() will try to simplify the result of lapply() if possible. Essentially, sapply() calls lapply()
on its input and then applies the following algorithm:  

  * If the result is a list where every element is length 1, then a vector is returned  
  * If the result is a list where every element is a vector of the same length (> 1), a matrix is returned.  
  * If it can’t figure things out, a list is returned  

```{r}
x<-list(a=matrix(1:4, 2,2), b=matrix(1:6,3,2)); x
sapply(x, mean)
library(datasets)
s <- split(airquality, airquality$Month)
sapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")],na.rm = TRUE)})
```
#### apply
The apply() function is used to a evaluate a function (often an anonymous one) over the margins
of an array. It is most often used to apply a function to the rows or columns of a matrix (which is
just a 2-dimensional array).
```{r}
str(apply)
```
Por ejemplo:
```{r}
x<-matrix(rnorm(200),20,10)
apply(x,2,mean) #dimension 2 is cols. El promedio se hace vertical
apply(x,1,sum)  #Sum through the rows 
```
Hay funciones como rowsum, o colsum que pueden reducir el codigo.   

#### mapply 
The mapply() function is a multivariate apply of sorts which applies a function in parallel over a set
of arguments. Recall that lapply() and friends only iterate over a single R object.
```{r}
str(mapply)
```
  *  FUN is a function to apply
  * ... contains arguments to apply over  
  * MoreArgs is a list of other arguments to FUN
  * SIMPLIFY indicates whether the result should be simplyfied  
  
Por ejemplo, en vez de usar `list(rep(1,4), rep(2,3), rep(3,2), rep(4,1))` se puede usar:
```{r}
mapply(rep,1:4,4:1)
```
  * `mapply()` can be used to vectorize a function:
```{r}
sumsq <- function(mu, sigma, x) {sum(((x - mu) / sigma)^2)}
x <- rnorm(100) ## Generate some data
sumsq(1:10, 1:10, x) #not the result that we want
mapply(sumsq, 1:10, 1:10, MoreArgs = list(x = x)) #We can use also vectorize
```
#### tapply
tapply() is used to apply a function over subsets of a vector. It can be thought of as a combination
of split() and sapply() for vectors only. I’ve been told that the “t” in tapply() refers to “table”,
but that is unconfirmed.  
```{r}
str(tapply)
```

  * X is a vector 
  * INDEX is a factor or a list of factors
  * FUN is a function to apply
  * ... contains arguments to apply over  
  * SIMPLIFY indicates whether the result should be simplyfied  
```{r}
x<-c(rnorm(10), runif(10),rnorm(10))
f<-gl(3,10) #Genera los niveles de los factores gl(n,k) n es el numero de niveles y k el numero de replicas
tapply(x,f,mean)
data("mtcars")
mtcars[1:5,]
tapply(mtcars$mpg,mtcars$cyl,mean)
```

#### split
_split_ takes a vector or other objetc and splits it into groups determined by a factor of list of factors
```
str(split)
function(x, f, drop=FALSE,... )
```
  * x is a vector or data frame  
  * f is a factores or a list of factors
  * drop indicates whether empty factors levels should be dropped

```{r}
x<-c(rnorm(10), runif(10),rnorm(10))
f<-gl(3,10) #Genera los niveles de los factores gl(n,k) n es el numero de niveles y k el numero de replicas
split(x,f) #siempre devuelve una lista. Se puede usar con otra function
lapply(split(x,f),mean)
```
  * **Usefull** example
```{r}
library(datasets)
s <- split(airquality, airquality$Month)
sapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")],na.rm = TRUE)})
```


### __*Regular Expressions*__ {.tabset .tabset-pills}
#### Intro
  * `grep()` `grepl()
  * `sapply()`
  * `apply()`
  * `mapply()` `vectorize()`
  * `tapply()`
  * `split()`  

#### Grep  
  
  
  
### __*Funciones creadas *__ {.tabset .tabset-pills}
#### Text-Telegram
  * Returns a text with START at the begining and STOP at the end
```{r}
telegram <- function(...){paste("START", ..., "STOP")}